Open-Set Domain Adaptation through Self-Supervision

### 2022-02-12 paper/main.tex
shorter bib

%------------------------------------------------------------------------
\section{Related Work}
\label{sec:relatedwork}

>>>>

%------------------------------------------------------------------------
\section{Related Work}
\label{sec:relatedwork}
\textbf{Closed-Set Domain Adaptation} is the setting where labeled training data is 
available on a source domain, but the goal is to have good performance on a 
unlabeled target domain with a different marginal distribution
of data. The aim is to align the learned representations of the source and target domains.

Methods for unsupervised domain adaptation in computer vision can be divided into 
three broad classes. The first class aims to induce alignment in some feature space optimizing the distributional discrepancy \cite{bousmalis2016domain,long2015learning,ganin2015unsupervised,hoffman2017cycada}.
The second class of methods directly transforms the source images to resemble 
the target images with generative models \cite{taigman2016unsupervised,tzeng2017adversarial,bousmalis2017unsupervised}, 
operating on image pixels directly instead of an intermediate 
representation space.
The third class uses a model trained on the labeled source data to 
estimate labels on the target data, then trains on some of those estimated 
pseudo-labels (e.g. the most confident ones), therefore bootstrapping through the 
unlabeled target data. This technique is borrowed from semi-supervised learning, 
where it is called co-training \cite{saito2017asymmetric,zou2018domain,chen2017rethinking}.

In contrast, the presented method uses a Self-Supervised approach, already presented in 
the recent years by many other works \cite{sun2019unsupervised,saito2020universa,yue2021prototypical}.
using pretext tasks (e.g., image rotation, jigsaw puzzle \cite{CarlucciJigsaw}, mutual information (MI) \cite{selfsupfeng}, instance discrimination \cite{chen2020simple}) to learn high-level feature representation in 
source and target domains by jointly training a shared feature extractor.

\textbf{Open Set Recognition} (OSR) has the objective of learning a classifier 
that can reject the unknown samples while classifying the known classes accurately.

The problem was first formulated in \cite{firstPaperOSR} and since then, several other works have analyzed this challenge in the context 
of deep networks \cite{perera2019deep, hein2019relu, bendale2015open}.
Some techinques presented in the past were focus on an adversarial approach to 
delineate closed from open-set images \cite{ge2017generative, kong2021opengan}.
Other approaches include reconstruction based methods 
\cite{yoshihashi2019classificationreconstruction}
which use poor test-time reconstruction as an open-set indicator.

In the recent years different works to a Self-Supervised solution, used also in the 
presented method, engaging in learning invariant representations to the 
transformations of the input data, improving separation of classes from each other 
and from open-set samples \cite{dhamija2021selfsupervised,jia2021selfsupervised,GeoTran}.

\textbf{Self-supervised Learning} is an emerging field focused on using 
data itself as supervision for auxiliary (also called “pretext”) 
tasks that learn deep feature representations, which will hopefully be informative 
for downstream “real” tasks.
Many such auxiliary tasks have been proposed in the literature, including
image colorization \cite{zhang2016colorful},
image jigsaw puzzle [\cite{noroozi2017unsupervised}, image inpainting \cite{pathak2016context},
contrastive learning \cite{oord2019representation}, geometric transformations \cite{dosovitskiy2015discriminative}.

Another geometric transformation task is rotation recognition \cite{OldROS}, where input images
are rotated by multiples of 90° and the network is trained to predict the rotation
angle pf each image. This pretext task has been successfully used for both anomaly
detection \cite{golan2018deep} and closed-set domain adaptation \cite{selfsupxu}.

\textbf{Open Set Domain Adaptation} (OSDA) is a challenging domain adaptation 
setting which allows the existence of unknown classes on the target domain.
The term "OSDA" was first introduced by Busto and Gall \cite{firstosda}, but the 
currently accepted definition was introduced by Saito \etal \cite{saito2018open} considering the target as 
containing all the source categories and additional set of private categories 
that should be considered unknown. They have presented Open Set
Back-Propagation (OSBP) \cite{saito2018open} as an adversarial method 
to increase the prediction variances so that the generator can choose if accept Ornot a target saple.
Separate To Adapt (STA) \cite{sat} train a multi-binary classifier to progressively 
separate the samples of unknown and known classes.
Universal Adaptation Network (UAN) \cite{uan}, originally proposed for the universal 
domain adaptation setting that is a superset of OSDA, quantify the sample-level 
transferability and recognize the unknown samples based on it.
Attract or Distract (AoD) \cite{feng2019attract} refine the decision by using 
metric learning to reduce the intra-class distance in known classes and push 
the unknown class away from the known classes.
Jing \etal \cite{Jing_Li_Zhu_Ding_Lu_Yang_2021} have recently proposed a method
to recognize the unknown samples according to the centroid deviation angles and
employing the statistical Extreme Value Theory to recognize the unknown 
samples that are misclassified into known classes.

The approach that we are going to present is instead based on a Self-Supervision 
task, the rotation recognition, used to both separate known and unknown target 
samples and align the known source and target distributions.
